---
title: "Chocolate Chip Cookies"
execute:
  error: true
author: "Your Name"
output: html_document
---

## Reading In the Data

First, read in the CSV data of cookie ingredients.
Make sure that your end-result data has appropriate types for each column - these should match the types provided in the documentation in the README.md file.

```{r}
cookie_data <- read.csv("choc_chip_cookie_ingredients.csv")

str(cookie_data)
```

```{python}
import pandas

df = pandas.read_csv("choc_chip_cookie_ingredients.csv")

print(df.dtypes)
```


## Exploratory Data Analysis

Exploratory data analysis is the process of getting familiar with your dataset. To get started, [this blog post](https://www.mrdbourke.com/a-gentle-introduction-to-exploratory-data-analysis/) provides a nice checklist to get you thinking:

> 1.  What question(s) are you trying to solve (or prove wrong)?
I am asking what my goal is and how I get there with problem solving.
> 2.  What kind of data do you have and how do you treat different types?
I have strings, integers, and floats, all of which I will treat differently and affect to bend my data.
> 3.  What's missing from the data and how do you deal with it?
Some data that is missing can be dropped if it is impossible to guess, but other data we can input the mean for a placeholder.
> 4.  Where are the outliers and why should you care about them?
Errors and statistically significant data points can be removed, other outliers will be kept to evaluate and see how the data changes.
> 5.  How can you add, change or remove features to get more out of your data?
Creating new columns or values for different things, and visualizing them differently with graphs or axis.

### Generating Questions

Generate at least 5 questions you might explore using this database of cookie ingredients.

1. What is the most common ingredient/oven temperature/amount?
2. Are there any strong relationships between any of the variables in the cookies?
3. What is the average number of units made or total amount of ingredients?
4. How do instructions vary?
5. Can we visualize why some cookies did better than others?


### Skimming the Data

One thing we often want to do during EDA is to examine the quality of the data - are there missing values? What quirks might exist in the dataset?

The `skimr` package in R, and the similar `skimpy` package in python (which has a much better name, in my opinion), can help provide visual summaries of the data. 

Install both packages, and read the package documentation ([R](https://cran.r-project.org/web/packages/skimr/vignettes/skimr.html), [Python](https://pypi.org/project/skimpy/).

[Part 1] Use each package and generate summaries of your data that require the use of at least some non-default options in each package's `skim` function.


```{r}
library (skimr)
library (readr)

package <- read.csv("choc_chip_cookie_ingredients.csv")

package_read <- skim(package)
seperate <- partition(package_read)
print(seperate)
yank(package_read, "numeric")
```

```{python}
from skimpy import skim
import pandas as pd

df = pd.read_csv("choc_chip_cookie_ingredients.csv")

cookies = skim(df)

numeric = cookies['numeric']
categorical = cookies['number']

print(numeric)
print(categorical)
```

[Part 2] Write 1-2 sentences about what you can tell from each summary display you generate. Did you discover anything new about the data?
Some data is singularly numeric and there are no "factor" data. It is also divided into decimals and whole number data sets.

### Generating Tables

Another useful technique for exploratory data analysis is to generate summary tables. 
You may want to use the `dplyr` package in R (`group_by` or `count` functions), as well as the `groupby` and `count` methods in Pandas. [Python example](https://sparkbyexamples.com/pandas/pandas-groupby-count-examples/), [R example](https://dplyr.tidyverse.org/reference/count.html)

[Part 1] Using R and Python, generate a table that shows what **proportion** of recipes contain each type of ingredient, for the most common 20 ingredients.
```{r}
library(dplyr)
library (readr)

cookies <- read.csv("choc_chip_cookie_ingredients.csv")

total_recipes <- length(unique(cookies$Recipe_Index))

ingredient_count <- cookies %>%
  group_by(Ingredient) %>%
  count() %>%
  arrange(desc(n)) %>%
  top_n(20) %>%
  mutate(proportion = n / total_recipes)  # Calculate the proportion of recipes containing each ingredient

print(ingredient_count)
```

```{python}
import pandas as pd

cookies = pd.read_csv("choc_chip_cookie_ingredients.csv")

total_recipes = cookies['Recipe_Index'].nunique()

ingredient_count = cookies.groupby('Ingredient').size().reset_index(name='n')
ingredient_count = ingredient_count.sort_values(by='n', ascending=False).head(20)
ingredient_count['proportion'] = ingredient_count['n'] / total_recipes

print(ingredient_count)
```
[Part 2] Print out a character string that lists all of the ingredients that do not appear in at least 20 recipes.
```{r}
ingredients_less_than_20 <- ingredient_count %>%
  filter(n < 20) %>%
  pull(Ingredient)

ingredients_string <- paste(ingredients_less_than_20, collapse = ", ")

print(ingredients_string)
```

```{python}
ingredients_less_than_20 = ingredient_count[ingredient_count['n'] < 20]['Ingredient']

ingredients_string = ', '.join(ingredients_less_than_20)

print(ingredients_string)
```

### Visualization

Using whatever plotting system you are comfortable with in R or python, see if you can create a couple of useful exploratory data visualizations which address one of the questions you wrote above - or another question which you've come up with as you've worked on this assignment.

[Part 1] Create at least one plot (it doesn't have to be pretty) that showcases an interesting facet of the data.
```{python}
import matplotlib.pyplot as plt

plt.figure(figsize=(10, 6))
plt.barh(ingredient_count['Ingredient'], ingredient_count['n'], color='skyblue')
plt.xlabel('Number of Recipes')
plt.ylabel('Ingredient')
plt.title('Top Ingredients in Chocolate Chip Cookie Recipes')
plt.gca().invert_yaxis() #Change display for the top
plt.show()
```
[Part 2] Write 2-3 sentences about what you can learn from that plot and what directions you might want to investigate from here.
This plot visually highlights the most common ingredients in the recipes, helping identify staples like flour, sugar, and chocolate chips. The next step might involve investigating whether there are correlations between ingredient usage and specific cookie types or even exploring how these ingredients relate to recipe difficulty or preparation time.